Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-01-08 main_web.tex
updates

A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
sensor was used to provide salinity and temperature measurements.  The
AUV is a powered untethered platform that operates at $1$-$3$ m/s in
the upper water column. It has a multicore GPU NVIDIA Jetson TX1
(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
was built on top of the autonomous Teleo-Reactive EXecutive
(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
that the measurements are conditionally independent because the
salinity is extracted from the conductivity sensor which is different
from the temperature sensor. We specify variance $0.25^2$ for both
errors. This is based on prior experience indicating that there can be some noise in temperature and salinity readings mostly because of dynamic water masses and uncertainty in AUV position. The initial survey transect was not sampled very densely so a reliable extraction of a nugget parameter from those data was not possible.

@@

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner.
The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data. In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect. But here the estimated nugget from the actual trial data is close to the specified level.

@@

Of course, the initial 
Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates that the correlation between the two variables might be slightly larger than the initially specified $0.5$ and that uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical and
theoretical curves are not very different, which indicates that the specified Gaussian
model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
after about $150$ m as was specified. The working assumption of a separable covariance
function is hence not unreasonable.

Overall, the spatial Gaussian model representation is of course a simplification of the changing water conditions, and even within the model formulation, the parameter specification could likely have been improved. 
However, we do not expect minor changes to have caused much difference in sampling paths.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS), Center of Excellence, project number
223254, Nansen Legacy Program, project number 276730, and the Applied Underwater Robotics Labortatory (AURLab). CT
and DG acknowledge support from the Swiss National Science Foundation,
project number 178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to
aknowledge support of Idiap Research Institute, his primary
affiliation in an early version of this manuscript. We thank the anonymous    reviewers for their careful reading of our manuscript and their many insightful comments and suggestions. The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and members of the NTNU AURLab for help with AUV deployments

>>>>

A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
sensor was used to provide salinity and temperature measurements.  The
AUV is a powered, untethered platform that operates at $1$-$3$ m/s in
the upper water column. It has a multicore GPU NVIDIA Jetson TX1
(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
was built on top of the autonomous Teleo-Reactive EXecutive
(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
that the measurements are conditionally independent because the
salinity is extracted from the conductivity sensor which is different
from the temperature sensor. We specify variance $0.25^2$ for both
errors. This is based on prior experience indicating that there can be
some noise in temperature and salinity readings mostly because of
dynamic water masses and uncertainty in AUV position. The initial
survey transect was not sampled very densely so a reliable extraction
of a nugget parameter from those data was not possible.

@@

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner.  The AUV paths in both runs are relatively
winding in the sense that the vehicle reacts to the data. In our
experience from simulated runs, the paths tend to be smoother when we
use a larger nugget effect. But here the estimated nugget from the
actual trial data is close to the specified level.

@@

Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity
and temperature are subtracted from the data. This scatter-plot of
joint residuals indicates that the correlation between the two
variables might be slightly larger than the initially specified $0.5$
and that uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
and theoretical curves are not very different, which indicates that
the specified Gaussian model fits reasonably
well. Fig.~\ref{fig:parest_c} shows the empirical variogram of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m as was
specified. The working assumption of a separable covariance function
is hence not unreasonable.

Overall, the spatial Gaussian model representation is a simplification
of the changing water conditions, and even within the model
formulation, the parameter specification could likely have been
improved.  However, we do not expect minor changes to have caused much
difference in sampling paths.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS), Center of Excellence, project number
223254, Nansen Legacy Program, project number 276730, and the Applied
Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
support from the Swiss National Science Foundation, project number
178858. JE and KR acknowledge support from Norwegian research council
(RCN), project number 305445. DG would like to aknowledge support of
Idiap Research Institute, his primary affiliation in an early version
of this manuscript. We thank the anonymous reviewers for their careful
reading of our manuscript and their many insightful comments and
suggestions. The authors would also like to thank Niklas Linde of the
University of Lausanne for providing constructive feedback about this
work, and members of the NTNU AURLab for help with AUV deployments.
